---
title: "ML CP"
author: "Chris De Coster"
date: "Sunday, October 26, 2014"
output: html_document
---
## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. 
These type of devices are part of the quantified self movement ? a group of enthusiasts who take measurements about themselves regularly to improve their health, 
to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, 
but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. 
They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 


## Data 
The training data for this project are available here: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

### Packages used
```{r message=FALSE, warning=FALSE}
library(caret)
library(randomForest)
library(data.table)
set.seed(1)
```

## Getting and Cleaning Data:
Files are dowmloaded and imported using read.csv
```{r message=FALSE, warning=FALSE}
if(!file.exists("pml-training.csv")){
  download.file(url="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile = "pml-training.csv")}
if(!file.exists("pml-testing.csv")){
  download.file(url="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = "pml-testing.csv")}

df<- read.csv(file = file.path(getwd(),'pml-training.csv'), na.strings=c('NA','','#DIV/0!'))
df_test <- read.csv(file = file.path(getwd(), 'pml-testing.csv'), na.strings=c('NA','','#DIV/0!')) 

```

Next, the columns with NA values are removed and also the first 7 columns because they do not contain data of any use for the prediction.
```{r message=FALSE, warning=FALSE}
na <- apply(df, 2, function(x) {sum(is.na(x))})
df <- subset(df[, which(na == 0)], select=-c(X, user_name, new_window, num_window, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp))
```

## Building the Model
### Split data:
I use 60% of the training data(file) to generate a training dataset and 40% for a testing dataset, partioned using the "classe " variable.

```{r message=FALSE, warning=FALSE}
inTrain <- createDataPartition(y = df$classe, p=0.60, list=FALSE)
training <- df[inTrain,]
testing <- df[-inTrain,]
```

### Prediction:
We train the dataset using the randowforest function of the caret package. The number of trees to grow is 20 to ensure that every input row gets predicted at least a few times.
```{r message=FALSE, warning=FALSE}
model <- train(training$classe ~., data = training, method="rf", ntree = 20)
prediction <- predict(model,newdata=training)
confusionMatrix(prediction, training$classe)
```
The matrix shows an accurancy of 1, this means no error. Very good, so there is no need to use another function or adapt parameters.

Now we target the testing dataset only once to get our final results:
```{r message=FALSE, warning=FALSE}
prediction <- predict(model,newdata=testing)
confusionMatrix(prediction, testing$classe)
```
This time the matrix shows a accurancy of 0.989, so an error of 0.011.

## Submission part:
Apply algorithm to test data:

```{r eval=FALSE}
na <- apply(df_test, 2, function(x) {sum(is.na(x))})
df_test <- subset(df_test[, which(na == 0)], select=-c(X, user_name, new_window, num_window, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp))

test_prediction <- predict(model, newdata = df_test)

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

# write the predictons generated in the previsou step into files.
pml_write_files(test_prediction)

```



